---
title: "Effective exchange rates"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Labour_cost}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.height = 7,
  fig.width = 10,
  fig.path = "Fig_eer/", 
  echo = FALSE, message = FALSE, warning = FALSE
)


# library(ficomp)

devtools::load_all()

library(dplyr)
library(forcats)
library(ggplot2)
library(eurostat)
library(tidyr)
library(ggptt)

set_proj_theme()

data("ert_eff_ic_m", "ert_eff_ic_q")

base_year <- 2008
base_years <- 2000:2018
geos <- c(Finland = "FI", Sweden = "SE", Germany = "DE", France = "FR")
start_time <- "1995-01-01"

```

# Eurostat - DG ECFIN

Double export weights, moving weights (t, t-1) 


* EA19= Euro-area Member States
* EU28 = EU Member States
* IC37 = EU28 + 9 other industrial countries (Australia, Canada, United States, Japan, Norway, New Zealand, Mexico, Switzerland and Turkey)
* Broad group (42) = IC37 + 5 other industrial countries (Russia, China, Brazil, South Korea and Hong Kong). For the broad group only HICP/CPI deflated REER are available.


### monthly data

```{r variables_m}

ert_eff_ic_m %>% 
  select(exch_rt, exch_rt_label) %>% 
  distinct() %>% 
  knitr::kable()

```

### quarterly data

```{r variables_q}

ert_eff_ic_q %>% 
  select(exch_rt, exch_rt_label) %>% 
  distinct() %>% 
  knitr::kable()

```

## Nominal effective exchange rates

```{r nominal}

ert_eff_ic_m %>% 
  filter(geo %in% geos,
         grepl("NEER", exch_rt)) %>% 
  ggplot(aes(time, values, colour = exch_rt_label)) +
  facet_wrap(~ geo) +
  geom_line() +
  the_legend_bot() +
  the_title_blank(c("x", "l")) +
  guides(colour = guide_legend(nrow = 4))


```

## Real effective exchange rates

### CPI mountly

```{r real_cpi_m}

ert_eff_ic_m %>% 
  filter(geo %in% geos,
         grepl("REER", exch_rt)) %>% 
  ggplot(aes(time, values, colour = exch_rt_label)) +
  facet_wrap(~ geo) +
  geom_line() +
  the_legend_bot() +
  the_title_blank(c("x", "l")) +
  guides(colour = guide_legend(nrow = 4))


```

### ULC quarterly

```{r real_ulc_q}

ert_eff_ic_q %>% 
  filter(geo %in% geos,
         grepl("ULCT", exch_rt)) %>% 
  ggplot(aes(time, values, colour = exch_rt_label)) +
  facet_wrap(~ geo) +
  geom_line() +
  the_legend_bot() +
  the_title_blank(c("x", "l")) +
  guides(colour = guide_legend(nrow = 4))


```

## Compined for 37 partners


```{r compined_37}

ert_eff_ic_q %>% 
  filter(grepl("ULCT", exch_rt)) %>% 
  bind_rows(ert_eff_ic_m) %>% 
  filter(geo %in% geos,
         grepl("IC37", exch_rt)) %>% 
  ggplot(aes(time, values, colour = exch_rt_label)) +
  facet_wrap(~ geo) +
  geom_line() +
  the_legend_bot() +
  the_title_blank(c("x", "l")) +
  guides(colour = guide_legend(nrow = 4))

```

```{r compined_37_geo}

ert_eff_ic_q %>% 
  filter(grepl("ULCT", exch_rt)) %>% 
  bind_rows(ert_eff_ic_m) %>% 
  filter(geo %in% geos,
         grepl("IC37", exch_rt)) %>% 
  ggplot(aes(time, values, colour = geo)) +
  facet_wrap(~ exch_rt) +
  geom_line() +
  the_legend_bot() +
  the_title_blank(c("x", "l")) 

```

```{r fi_all37_q}

ert_eff_ic_q %>% 
  filter(geo == "FI" ,
         grepl("IC37", exch_rt)) %>% 
  ggplot(aes(time, values, colour = exch_rt_label)) +
  facet_wrap(~ geo) +
  geom_line() +
  the_legend_bot() +
  the_title_blank(c("x", "l")) +
  guides(colour = guide_legend(nrow = 4))


```
