---
title: "Kilpailukykymuuttujat"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Labour_cost}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.height = 6,
  fig.width = 8,
  fig.path = "Report/", 
  echo = FALSE, message = FALSE, warning = FALSE
)


# library(ficomp)

devtools::load_all()

library(dplyr)
library(forcats)
library(ggplot2)
library(eurostat)
library(tidyr)
library(ggptt)

set_proj_theme()

data(ulc_oecd_dat, q_dat, a_dat, naq_eurostat_dat, eo_q_dat)

base_year <- 2010
base_years <- 2000:2018
geos <- c(Finland = "FI", Sweden = "SE", Germany = "DE", France = "FR")
start_time <- "1996-01-01"
geo_trans <- geo_fi
h_geos <- c("FI", "SE", "DE", "US")


```

# Data

Data for unit labour cost index

```{r}

tibble(code = levels(naq_eurostat_dat$na_item), name = label_eurostat(levels(naq_eurostat_dat$na_item), "na_item"))

```


# Nimellinen yksikkötyökustannus

## OECD quarterly ULC data

OECD ULC tiedoissa olevat maat: `r levels(ulc_oecd_dat$geo)`. 

Maiden lukumäärä `r length(levels(ulc_oecd_dat$geo))`.

Muuttujat: `r levels(ulc_oecd_dat$na_item)`


```{r oecd_ulc_q_all}



ulc_oecd_dat %>% 
  filter(na_item == "NULC_APER") %>% 
  mutate(h_geo = fct_other(geo, keep = h_geos,other_level = "muut"),
         h_geo = fct_relevel(h_geo, h_geos),
         geo = fct_relevel(geo, rev(h_geos), after = Inf),
         size = !(geo %in% h_geo)) %>% 
  translate(h_geo, geo_trans) %>% 
  ggplot(aes(time, values, group = geo, colour = h_geo, size = size)) +
  geom_line() +
  scale_size_manual(values = c(2,1), guide = "none") +
  scale_colour_manual(values = geo_col(length(h_geos) + 1)) +
  the_title_blank()

# library(ggiraph)
# 
# ggi <- last_plot() + geom_line_interactive(aes(tooltip = geo))
# girafe(print(ggi))

```

Tiedot alkavat pääosin vuodesta 1995. Aloittamalla vuodesta 1996 saadaan muutama muu maa mukaan.
Tietojen saatavuuden ja vertailtavuuden perusteella mukaan on otettu seuraavat maat.

```{r oecd_ulc_q_stat, fig.height=7}

ulc_oecd_dat %>% 
  filter(na_item == "NULC_APER") %>% 
  drop_na(values) %>% 
  group_by(geo, na_item) %>% 
  mutate(tmin = min(time), 
         tmax = max(time)) %>% 
  ungroup() %>% 
  mutate(selected = geo %in% c(eurostat_geos, oecd_geos_ulcq)) %>% 
  translate(geo, geo_trans) %>% 
  ggplot(aes(geo, ymin = tmin, ymax = tmax, colour = selected)) +
    geom_linerange(size = 2) +
    coord_flip() +
    geom_hline(yintercept = as.numeric(as.Date("1996-01-01"))) +
    scale_y_date(breaks = as.Date(c("1960-01-01", "1980-01-01", 
                                    "1996-01-01", "2000-01-01", "2020-01-01")), 
                 date_labels = "%Y") +
    scale_colour_manual(values = rev(geo_col(2)), guide = "none") +
    the_title_blank("y")

save_fig("2_ulc_oecd_selected_geo", height = 5.8)

```


### Valitut maat OECD quarterly

Indeksi ja suhteellinen indeksi IMF painoilla. y-akseli on katkaistu. 

```{r oecd_ulc_q_selected}



q_dat_oecd_ulc %>% 
  # filter(geo != "JP") %>%
  select(geo, time, nulc_aper, nulc_aper_rel_imf) %>% 
  gather(vars, values, nulc_aper, nulc_aper_rel_imf) %>% 
  mutate(h_geo = fct_other(geo, keep = h_geos,other_level = "muut"),
         h_geo = fct_relevel(h_geo, h_geos),
         geo = fct_relevel(geo, rev(h_geos), after = Inf),
         size = !(geo %in% h_geo)) %>% 
  translate(h_geo, geo_trans) %>% 
  translate(vars, c(nulc_aper = "Indeksi", 
                    nulc_aper_rel_imf = "Suhteellinen indeksi")) %>% 
  ggplot(aes(time, values, group = geo, colour = h_geo, size = size)) +
  facet_wrap(~ vars) +
  geom_line() +
  scale_size_manual(values = c(2,1), guide = "none") +
  scale_colour_manual(values = geo_col(length(h_geos) + 1)) +
  the_title_blank(c("x", "l")) +
  scale_y_continuous(limits = c(70,130), expand = expansion(mult = 0, add = 0)) +
  labs(y = "indeksi, 2010 = 100")

save_fig("3_ulc_oecd")

```

# OECD Economic Outlook

Alkaa 1991, mutta neljännesvuositieodoissa osasta maista ei tietoja. c("NO", "DK", "NZ", "ES")

```{r oecd_eo_q_stat}

eo_q_dat_large %>% 
  select(geo, time, values = nulc) %>% 
  group_by(geo) %>% 
  # coaleasce and na_if to set same min and ma for countries with all missing
  mutate(tmin = coalesce(na_if(min(time[!is.na(values)]), Inf), as.Date("1996-01-01")), 
         tmax = coalesce(na_if(max(time[!is.na(values)]), -Inf), as.Date("1996-01-01"))) %>% 
  ungroup() %>% 
  mutate(selected = geo %in% c(eurostat_geos, oecd_geos_ulcq)) %>% 
  translate(geo, geo_trans) %>% 
  ggplot(aes(geo, ymin = tmin, ymax = tmax, colour = selected)) +
    geom_linerange(size = 2, colour = geo_col(2)[1]) +
    coord_flip() +
    geom_hline(yintercept = as.numeric(as.Date("1991-01-01"))) +
    scale_y_date(breaks = as.Date(c("1960-01-01", "1980-01-01", 
                                    "1991-01-01", "2000-01-01", "2020-01-01")), 
                 date_labels = "%Y") +
    scale_colour_manual(values = rev(geo_col(2)), guide = "none") +
    the_title_blank("y")



```

# Neljännesvuosidata

Yhdistelmä Eurostatin ja OECD:n Quarterly National Accounts datasta 


```{r q_dat_stat}

q_dat %>% 
  select(geo, time, values = nulc_aper) %>% 
  group_by(geo) %>% 
  mutate(tmin = min(time[!is.na(values)]), 
         tmax = max(time[!is.na(values)])) %>% 
  ungroup() %>% 
  mutate(selected = geo %in% c(eurostat_geos, oecd_geos_ulcq)) %>% 
  translate(geo, geo_trans) %>% 
  ggplot(aes(geo, ymin = tmin, ymax = tmax, colour = selected)) +
    geom_linerange(size = 2, colour = geo_col(2)[1]) +
    coord_flip() +
    geom_hline(yintercept = as.numeric(as.Date("1996-01-01"))) +
    scale_y_date(breaks = as.Date(c("1996-01-01", "2000-01-01",
                                    "2020-01-01")),
    date_labels = "%Y") +
    scale_colour_manual(values = rev(geo_col(2)), guide = "none") +
    the_title_blank("y")


```

```{r q_dat_nulc}

q_dat %>% 
  # filter(geo != "JP") %>%
  select(geo, time, nulc_aper, nulc_aper_rel_imf) %>% 
  gather(vars, values, nulc_aper, nulc_aper_rel_imf) %>% 
  mutate(h_geo = fct_other(geo, keep = h_geos,other_level = "muut"),
         h_geo = fct_relevel(h_geo, h_geos),
         geo = fct_relevel(geo, rev(h_geos), after = Inf),
         size = !(geo %in% h_geo)) %>% 
  translate(h_geo, geo_trans) %>% 
  translate(vars, c(nulc_aper = "Indeksi", 
                    nulc_aper_rel_imf = "Suhteellinen indeksi")) %>% 
  ggplot(aes(time, values, group = geo, colour = h_geo, size = size)) +
  facet_wrap(~ vars) +
  geom_line() +
  scale_size_manual(values = c(2,1), guide = "none") +
  scale_colour_manual(values = geo_col(length(h_geos) + 1)) +
  the_title_blank(c("x", "l")) +
  scale_y_continuous(limits = c(70,130), expand = expansion(mult = 0, add = 0)) +
  labs(y = "indeksi, 2010 = 100")

```

## Vertailu OECD ULC ja yhdistelmä

```{r comb_fi}

bind_rows("Eurostat ja OECD QNA" = q_dat,
          "OECD ULC" = q_dat_oecd_ulc,
          "OECD EO" = rename(eo_q_dat, nulc_aper = nulc, nulc_aper_rel_imf = nulc_rel_imf),
          .id = "source") %>% 
  filter(geo == "FI", time >= start_time) %>% 
  select(source, time, nulc_aper, nulc_aper_rel_imf) %>% 
  gather(vars, values, nulc_aper, nulc_aper_rel_imf) %>% 
  # mutate(h_geo = fct_other(geo, keep = h_geos,other_level = "muut"),
  #        h_geo = fct_relevel(h_geo, h_geos),
  #        geo = fct_relevel(geo, rev(h_geos), after = Inf),
  #        size = !(geo %in% h_geo)) %>% 
  # translate(h_geo, geo_trans) %>% 
  translate(vars, c(nulc_aper = "Indeksi", 
                    nulc_aper_rel_imf = "Suhteellinen indeksi")) %>% 
  ggplot(aes(time, values, colour = source)) +
  facet_wrap(~ vars) +
  geom_line() +
  # scale_size_manual(values = c(2,1), guide = "none") +
  # scale_colour_manual(values = geo_col(length(h_geos) + 1)) +
  the_title_blank(c("x", "l")) +
  # scale_y_continuous(limits = c(70,130), expand = expansion(mult = 0, add = 0)) +
  labs(y = "indeksi, 2010 = 100")
  


```

```{r comb_all, fig.height=15}

bind_rows("Eurostat ja EO" = q_dat,
          "OECD ULC" = q_dat_oecd_ulc, .id = "source") %>% 
  select(geo, source, time, nulc_aper, nulc_aper_rel_imf) %>% 
  gather(vars, values, nulc_aper, nulc_aper_rel_imf) %>% 
  # mutate(h_geo = fct_other(geo, keep = h_geos,other_level = "muut"),
  #        h_geo = fct_relevel(h_geo, h_geos),
  #        geo = fct_relevel(geo, rev(h_geos), after = Inf),
  #        size = !(geo %in% h_geo)) %>% 
  # translate(h_geo, geo_trans) %>% 
  mutate(source = fct_rev(as_factor(source))) %>% 
  translate(vars, c(nulc_aper = "Indeksi", 
                    nulc_aper_rel_imf = "Suhteellinen indeksi")) %>% 
  ggplot(aes(time, values, colour = source)) +
  facet_grid(geo ~ vars) +
  geom_line() +
  # scale_size_manual(values = c(2,1), guide = "none") +
  # scale_colour_manual(values = geo_col(length(h_geos) + 1)) +
  the_title_blank(c("x", "l")) +
  # scale_y_continuous(limits = c(70,130), expand = expansion(mult = 0, add = 0)) +
  labs(y = "indeksi, 2010 = 100")
  
```

## Different measures quarterly

Different definations for unit labour cost:
* Output: GDP or value added
* Relation to hours worked or persons (relevant if adjusted for self-employed or for decomposition)

```{r nulc}

lc_dat_nace <- 
  naq_eurostat_dat %>% 
  unite(vars, na_item, unit) %>%  
  # filter(vars %in% c("D1_CP_MNAC", "B1G_CLV10_MNAC")) %>% 
  spread(vars, values) %>% 
  group_by(geo, nace_r2) %>% 
  mutate(
    # nominal unit labour cost
    nulc_va = ind_ulc(D1_CP_MNAC, B1G_CLV10_MNAC, time = time, baseyear = base_year),
    nulc = ind_ulc(D1_CP_MNAC, B1GQ_CLV10_MNAC, time = time, baseyear = base_year),
    # nominal unit labour cost, adjusted to take account emploees/emplyed share
    nulc_hw_va = ind_ulc(D1_CP_MNAC / SAL_DC_THS_HW, B1G_CLV10_MNAC / EMP_DC_THS_HW, time = time, baseyear = base_year),
    nulc_hw = ind_ulc(D1_CP_MNAC / SAL_DC_THS_HW, B1GQ_CLV10_MNAC / EMP_DC_THS_HW, time = time, baseyear = base_year),
    nulc_hw_eur = ind_ulc(D1_CP_MEUR / SAL_DC_THS_HW, B1GQ_CLV10_MNAC / EMP_DC_THS_HW, time = time, baseyear = base_year),
    nulc_per_va = ind_ulc(D1_CP_MNAC / SAL_DC_THS_PER, B1G_CLV10_MNAC / EMP_DC_THS_PER, time = time, baseyear = base_year),
    nulc_per = ind_ulc(D1_CP_MNAC / SAL_DC_THS_PER, B1GQ_CLV10_MNAC / EMP_DC_THS_PER, time = time, baseyear = base_year)) %>% 
  ungroup()

lc_dat <- lc_dat_nace %>% 
  filter(nace_r2 == "TOTAL")

lc_dat %>% 
  filter(geo %in% geos,
         time >= start_time) %>% 
  select(geo, time, nulc, nulc_va, nulc_hw, nulc_hw_va, nulc_per, nulc_per_va) %>% 
  gather(ind, values, -geo, -time) %>% 
  ggplot(aes(time, values, colour = geo)) +
  geom_line() +
  facet_wrap(~ ind) +
  the_title_blank() 

```


On short term a NULC indicator selection does not matter. On longer term
differences acculumate. 

```{r nulc_fi}

lc_dat %>% 
  filter(geo == "FI",
         time >= start_time) %>% 
  select(geo, time, nulc, nulc_va, nulc_hw, nulc_hw_va, nulc_per, nulc_per_va) %>% 
  gather(ind, values, -geo, -time) %>% 
  group_by(geo, ind) %>% 
  mutate(change = values / lag(values, 4, order_by = time) - 1) %>% 
  ungroup() %>% 
  gather(vars, values, values, change) %>% 
  ggplot(aes(time, values, colour = ind)) +
  facet_wrap(~ vars, scales = "free_y") +
  geom_line() +
  the_title_blank()

```



## ULC yearly from stan

```{r nulc_stan}

lc_dat_stan <- 
  stan_dat %>% 
  # filter(vars %in% c("D1_CP_MNAC", "B1G_CLV10_MNAC")) %>% 
  group_by(geo, nace_r2) %>% 
  mutate(
    # nominal unit labour cost, adjusted to take account emploees/emplyed share
    nulc_hw_va = ind_ulc(D1__CP_MNAC / SAL_DC__THS_HW, B1G__CLV10_MNAC / EMP_DC__THS_HW, time = time, baseyear = base_year),
    nulc_va = ind_ulc(D1__CP_MNAC, B1G__CLV10_MNAC, time = time, baseyear = base_year),
) %>% 
  ungroup()

lc_dat_stan %>% 
   filter(geo %in% c(geos, "US"),
          nace_r2 == "TOTAL") %>% 
  select(geo, time, nulc_va, nulc_hw_va) %>% 
  gather(ind, values, -geo, -time) %>% 
  ggplot(aes(time, values, colour = geo)) +
  geom_line() +
  facet_wrap(~ ind)

```

### In own and in common currency

```{r nulc_eur}

lc_dat %>% 
  filter(geo %in% geos,
         time >= start_time) %>% 
  select(geo, time, nulc_hw, nulc_hw_eur) %>% 
  gather(ind, values, -geo, -time) %>% 
  ggplot(aes(time, values, colour = geo)) +
  geom_line() +
  facet_wrap(~ ind) +
  the_title_blank()

```


```{r nulc_per}

lc_dat %>% 
  filter(geo %in% geos,
         time >= start_time) %>% 
  select(geo, time, nulc_per) %>% 
  gather(ind, values, -geo, -time) %>% 
  mutate(geo = fct_recode(geo, !!!geos)) %>% 
  ggplot(aes(time, values, colour = geo)) +
  geom_line() +
  labs(title = "Unit labour cost, employment based",
       y = "index, 2008 = 100") +
  the_title_blank(c("x", "l"))

```




# Nominal unit labour cost comparison from different sources

NULC_APER_OECD = NULC_PER_ECB = nulc_per 
  = (D1_CP_MNAC / SAL_DC_THS_PER) / (B1GQ_CLV10_MNAC / EMP_DC_THS_PER)
  
NULC_HW_ECB = nulc_hw 
  = (D1_CP_MNAC / SAL_DC_THS_HW) / (B1GQ_CLV10_MNAC / EMP_DC_THS_HW)

```{r nulc_comp}

ulc_oecd_with <- ulc_oecd_dat %>% 
  select(- unit) %>% 
  filter(na_item == "NULC_APER") %>% 
  mutate(na_item = paste0(na_item, "_OECD")) %>% 
  group_by(geo, na_item) %>% 
  mutate(values = rebase(values, time, base_year)) %>% 
  ungroup() %>% 
  spread(na_item, values)

ulc_eurostat_with <- ulc_eurostat_dat %>% 
  filter(na_item %in% c("NULC_HW", "NULC_PER")) %>% 
  select(time, geo, na_item, values) %>% 
  mutate(na_item = paste0(na_item, "_eurostat")) %>% 
  group_by(geo, na_item) %>% 
  mutate(values = rebase(values, time, base_year)) %>% 
  ungroup() %>% 
  spread(na_item, values)

ulc_ecb_with <- ulc_ecb_dat %>% 
  filter(na_item %in% c("NULC_HW", "NULC_PER")) %>% 
  select(time, geo, na_item, values) %>% 
  mutate(na_item = paste0(na_item, "_ECB")) %>% 
  group_by(geo, na_item) %>% 
  mutate(values = rebase(values, time, base_year)) %>% 
  ungroup() %>% 
  spread(na_item, values)

lc_with <- lc_dat %>% 
  select(time, geo, nulc, nulc_va, nulc_hw, nulc_hw_va, nulc_per, nulc_per_va) %>% 
  left_join(ulc_oecd_with, by = c("time", "geo")) %>% 
  # left_join(ulc_eurostat_with, by = c("time", "geo")) %>%
  left_join(ulc_ecb_with, by = c("time", "geo")) # %>%   filter(geo == "FI") %>% View()

lc_with %>% 
  gather(vars, values, - time, - geo) %>% 
  mutate(vars = fct_relevel(vars, "NULC_APER", after = Inf)) %>% 
  filter(geo == "FI") %>% 
  ggplot(aes(time, values, colour = vars)) +
  geom_line()


```





# Nominal unit labour cost manufacturing

```{r ulc_c}

lc_dat_c <- 
  naq_eurostat_dat %>% 
  filter(nace_r2 == "C") %>% 
  unite(vars, na_item, unit) %>% 
  # filter(vars %in% c("D1_CP_MNAC", "B1G_CLV10_MNAC")) %>% 
  spread(vars, values) %>% 
  group_by(geo) %>% 
  mutate(
    # nominal unit labour cost
    ulc = ind_ulc(D1_CP_MNAC, B1G_CLV10_MNAC, time = time, baseyear = base_years),
    # nominal unit labour cost, adjusted to take account emploees/emplyed share
    ulc_adj = ind_ulc(D1_CP_MNAC / SAL_DC_THS_HW, B1G_CLV10_MNAC / EMP_DC_THS_HW, time = time, baseyear = base_years)) %>% 
  ungroup()

lc_dat_c %>% 
  filter(geo %in% geos,
         time >= start_time) %>% 
  select(geo, time, ulc, ulc_adj) %>% 
  gather(ind, values, -geo, -time) %>% 
  ggplot(aes(time, values, colour = geo)) +
  geom_line() +
  facet_wrap(~ ind)

```

# Hourly compensation

```{r comp_h}


naq_eurostat_dat %>% 
  filter(nace_r2 %in% c("TOTAL", "C")) %>% 
  unite(vars, na_item, unit) %>% 
  # filter(vars %in% c("D1_CP_MNAC", "B1G_CLV10_MNAC")) %>% 
  spread(vars, values) %>% 
  group_by(geo, nace_r2) %>% 
  mutate(comp_h = rebase(D1_CP_MNAC / SAL_DC_THS_HW, time = time, baseyear = base_year),
         d_comp_h = 100 * (comp_h / lag(comp_h, 4) - 1)) %>% 
  ungroup() %>% 
  filter(geo %in% geos,
         time >= start_time) %>% 
  ggplot(aes(time, d_comp_h, colour = geo)) +
  geom_line() +
  facet_wrap(~ nace_r2)

```
```{r d_comp_h}


naq_eurostat_dat %>% 
  filter(nace_r2 %in% c("TOTAL", "C")) %>% 
  unite(vars, na_item, unit) %>% 
  # filter(vars %in% c("D1_CP_MNAC", "B1G_CLV10_MNAC")) %>% 
  spread(vars, values) %>% 
  group_by(geo, nace_r2) %>% 
  mutate(comp_h = rebase(D1_CP_MNAC / SAL_DC_THS_HW, time = time, baseyear = base_year),
         d_comp_h = 100 * (comp_h / lag(comp_h, 4) - 1)) %>% 
  ungroup() %>% 
  filter(geo %in% c("FI", "SE"),
         time >= "2010-01-01") %>% 
  ggplot(aes(time, d_comp_h, colour = geo)) +
  geom_line() +
  facet_wrap(~ nace_r2)

```
```{r q_dat_geos}

unique(q_dat$geo)

```


```{r q_dat}

h_geos <- "FI"

q_dat %>% 
  select(geo, time, nulc_aper, nulc_aper_rel15) %>% 
  gather(vars, values, -geo, -time) %>% 
  mutate(h_geo = fct_other(geo, keep = h_geos,other_level = "muut"),
         geo = fct_relevel(geo, h_geos, after = Inf)) %>% 
  ggplot(aes(time, values, group = geo, colour = h_geo, size = h_geo)) +
  geom_line() +
  facet_wrap(~ vars, scales = "free") +
  scale_size_manual(values = c(2,1)) +
  the_title_blank()

```

```{r q_dat_weighting}


q_dat %>% 
  filter(geo %in% c("FI", "DE", "FR", "IT")) %>% 
  select(geo, time, nulc_aper_rel15, nulc_aper_rel_bis) %>% 
  gather(vars, values, -geo, -time) %>% 
  # mutate(h_geo = fct_other(geo, keep = h_geos,other_level = "muut"),
  #        geo = fct_relevel(geo, h_geos, after = Inf)) %>% 
  ggplot(aes(time, values, colour = vars)) +
  geom_line() +
  facet_wrap(~ geo, scales = "free") +
  scale_size_manual(values = c(2,1)) +
  the_title_blank()

```

```{r q_dat_ind}


q_dat %>% 
  filter(geo %in% c("FI", "DE", "FR", "IT")) %>% 
  select(geo, time, nulc_aper_rel15, gdp_ind_rel15, exp_ind_rel15) %>% 
  mutate(inv_nulc_aper_rel15 = rebase(1/nulc_aper_rel15, time, base_year)) %>% 
  select(-nulc_aper_rel15) %>% 
  gather(vars, values, -geo, -time) %>% 
  # mutate(h_geo = fct_other(geo, keep = h_geos,other_level = "muut"),
  #        geo = fct_relevel(geo, h_geos, after = Inf)) %>% 
  ggplot(aes(time, values, colour = vars)) +
  geom_line() +
  facet_wrap(~ geo, scales = "free") +
  scale_size_manual(values = c(2,1)) +
  the_title_blank()

```

```{r a_dat_geos}

unique(a_dat$geo)

```

```{r a_dat}

h_geos <- "FI"

a_dat %>% 
  select(-contains("__")) %>% 
  filter(nace_r2 == "TOTAL") %>% 
  select(-nace_r2) %>% 
  gather(vars, values, -geo, -time) %>% 
  mutate(h_geo = fct_other(geo, keep = h_geos,other_level = "muut"),
         geo = fct_relevel(geo, h_geos, after = Inf)) %>% 
  ggplot(aes(time, values, group = geo, colour = h_geo, size = h_geo)) +
  geom_line() +
  facet_wrap(~ vars, scales = "free") +
  scale_size_manual(values = c(2,1)) +
  the_title_blank()

```

```{r a_dat_ind}


a_dat %>% 
  filter(nace_r2 == "TOTAL") %>% 
  select(-nace_r2) %>% 
  filter(geo %in% c("FI", "DE", "FR", "IT")) %>% 
  select(geo, time, nulc_hw_va_rel15, rulc_hw_va_rel15, gdp_ind_rel15, exp_ind_rel15) %>% 
  mutate(inv_nulc_hw_va_rel15 = rebase(1/nulc_hw_va_rel15, time, base_year),
         inv_rulc_hw_va_rel15 = rebase(1/rulc_hw_va_rel15, time, base_year)) %>% 
  select(-nulc_hw_va_rel15, -rulc_hw_va_rel15) %>% 
  gather(vars, values, -geo, -time) %>% 
  # mutate(h_geo = fct_other(geo, keep = h_geos,other_level = "muut"),
  #        geo = fct_relevel(geo, h_geos, after = Inf)) %>% 
  ggplot(aes(time, values, colour = vars)) +
  geom_line() +
  facet_wrap(~ geo, scales = "free_y") +
  scale_size_manual(values = c(2,1)) +
  the_title_blank()
```

```{r a_dat_ind_eur}


a_dat %>% 
  filter(nace_r2 == "TOTAL") %>% 
  select(-nace_r2) %>% 
  filter(geo %in% c("FI", "DE", "FR", "IT")) %>% 
  select(geo, time, nulc_hw_va_rel15, nulc_hw_va_eur_rel15, gdp_ind_rel15, exp_ind_rel15) %>% 
  mutate(inv_nulc_hw_va_eur_rel15 = rebase(1/nulc_hw_va_eur_rel15, time, base_year),
         inv_nulc_hw_va_rel15 = rebase(1/nulc_hw_va_rel15, time, base_year)) %>% 
  select(-nulc_hw_va_eur_rel15, -nulc_hw_va_rel15) %>% 
  gather(vars, values, -geo, -time) %>% 
  # mutate(h_geo = fct_other(geo, keep = h_geos,other_level = "muut"),
  #        geo = fct_relevel(geo, h_geos, after = Inf)) %>% 
  ggplot(aes(time, values, colour = vars)) +
  geom_line() +
  facet_wrap(~ geo, scales = "free_y") +
  scale_size_manual(values = c(2,1)) +
  the_title_blank()

```
